Modelling last glacial cycle ice dynamics in the Alps

### 2018-07-21 articles/cycle/response-referee2.tex

% ----------------------------------------------------------------------

\sechead{Specific comments}

    \referee{%
        I find missing something that indicates that the times are before
        present, as in line 5, line 8 and line 16 on page 1 -- and elsewhere in
        the paper it is written (120-0 ka) should you add before present or BP
        to indicate the time interval?}

    We have now defined this better in the introduction:

    \msquote{
        During the last 800\,ka (thousand years before present)}

@@

    \referee{%
        Could the reason for too large ice volumes when using the GRIP record,
        as mentioned in lines 13-15 on page 9 be due to Arctic Amplification?
        Could that have an effect then as it has now? This is also mentioned in
        line 11 on page 10.}

    These passages refer to the Younger Dryas and Dansgaard-Oeschger events.
    One must keep in mind that a linear scaling was applied to the GRIP record,
    so that it is already ``de-amplified'' before being used as temperature
    forcing.  Arctic amplification could be an explanation if it was stronger
    then than for changes that lead the LGM climate. Although we lack expertise
    on this topic, we could not find literature supporting this idea. Instead,
    current theories for the Younger Dryas and Dansgaard-Oeschger events mostly
    involve large amounts of freshwater discharge into the North Atlantic
    implying a more regional mechanism than arctic amplification. Therefore, we
    decided not mention arctic amplification in the text.

### 2018-07-22 articles/cycle/response-referee3.tex

% ----------------------------------------------------------------------

\sechead{Introduction}

    \referee{\textbf{p.~2, l.~28:}
        Ballantyne and Stone (2015) should be added to this list.}

    This reference is indeed relevant to the discussion on trimlines. We have
    added the reference in the introduction and in the corresponding section in
    the discussion.

@@

    \referee{\textbf{p.~2, l.~34:}
        It should be stated that it could be a consequence of both
        glaioclimatic interactions and uncertainties in dating methods.}

    We have added ``or both''.

    \referee{\textbf{p.~3, l.~2:}
        These points serve the literature well to highlight gaps for future
        research. However, I would argue that you do not get very far here on 1
        and 5 and do not completely solve the other 3 points. Your text
        reflects these shortcomings very well, for which you should be
        applauded. Though I think at this stage of the manuscript, your
        statement of intent, you should state that you do not claim to solve
        these questions, but rather push forward on all of them using your new
        approach of ice sheet modelling.}

    In order to clarify our statement of intent, the following text was added:

### 2018-07-22 articles/cycle/response-referee3.tex

% ----------------------------------------------------------------------

\sechead{Section 2.6}

    \referee{\textbf{p.~7:}
        The spatial distribution of your modern climate variables (precip,
        temp) will be massively influenced by elevation. Though there is a
        lapse rate, does this pattern of high precip and low temp over
        mountains remain throughout the simulation despite ice surface
        topography, and if so, how does this influence your results?}

    The temperature lapse rate corrections dampen local high temperatures in
    ice-filled valleys. The precipitation pattern, on the other hand, remains
    the same throughout the simulation, and this results in higher accumulation
    above the location of present-day mountains even if they are buried by ice
    in the model.

    Although we expect that small-scale variations in surface mass balance are
    smoothed out by ice flow, the presence of an ice sheet may have
    redistributed precipitation on a larger scale, for instance from the inner
    ranges to the foothills, or inequally between north and south of the Alps,
    etc. However, global precipitation reductions appear to have little effect
    on the average ice thickness (Fig.~2b, compare dark and light curves).
    Nevertheless, the new text shortly mentions this caveat:

    \msquote{%
        Besides [circulation changes], changes in the ice surface topography
        may have redistributed orographic precipitation.}

    Our conclusions also highlight the need for a more realistic forcing.

    \msquote{%
        Using [...] a more realistic climate forcing based on regional
        circulation model output [...] future modelling studies will certainly
        be able to quantify uncertainties associated with some of the above
        limitations.}

### 2018-07-22 articles/cycle/alps-cycle.tex
Address referee 3 comment on precipitation change.

    Finally, in some simulations (hereafter labelled PP), precipitation was
    reduced with air temperature in
    order to simulate the potential rarefaction of atmospheric moisture in
    colder climates. This was done using an empirical relationship derived from
    observed accumulation rates and oxygen isotopes concentrations in the GRIP
    ice core \citep{Dahl-Jensen.etal.1993},
    %
    \begin{align}
      {\Psi}_{\text{PP}}(t) = \exp[\psi{\Delta}T_{\text{TS}}(t)] \,,
    \end{align}
    %
    with $\psi=0.169/2.4=0.0704$ \citep{Huybrechts.2002}. This simple
    relationship likely does not reflect the complexity of atmospheric
    circulation changes that governed moisture availability over the Alps
    during the last glacial cycle. Palaeoclimate proxies indicate slightly
    reduced LGM precipitation in western Europe with anomalies diminishing
    eastwards \citep{Wu.etal.2007}. Regional circulation models indicate
    generally dryer conditions during MIS~3 \citep{Barron.Pollard.2002,
    Kjellstrom.etal.2010} but more precipitation south of the Alps during MIS~2
    LGM \citep{Strandberg.etal.2011, Ludwig.etal.2016}.
    Thus, another set of simulations presented here use constant
    precipitation, corresponding to $\psi=0$.

>>>>

    Finally, in some simulations (hereafter labelled PP), precipitation was
    reduced with air temperature in
    order to simulate the potential rarefaction of atmospheric moisture in
    colder climates. This was done using an empirical relationship derived from
    observed accumulation rates and oxygen isotopes concentrations in the GRIP
    ice core \citep{Dahl-Jensen.etal.1993},
    %
    \begin{align}
      {\Psi}_{\text{PP}}(t) = \exp[\psi{\Delta}T_{\text{TS}}(t)] \,,
    \end{align}
    %
    with $\psi=0.169/2.4=0.0704$ \citep{Huybrechts.2002}. The control set of
    simulations use constant precipitation, corresponding to $\psi=0$.
    This simple relationship does not reflect the complexity of atmospheric
    circulation changes that governed moisture availability over the Alps
    during the last glacial cycle. Palaeoclimate proxies indicate slightly
    reduced LGM precipitation in western Europe with anomalies diminishing
    eastwards \citep{Wu.etal.2007}. Regional circulation models indicate
    generally dryer conditions during MIS~3 \citep{Barron.Pollard.2002,
    Kjellstrom.etal.2010} but more precipitation south of the Alps during MIS~2
    LGM \citep{Strandberg.etal.2011, Ludwig.etal.2016}. Besides, changes in the
    ice surface topography may have redistributed orographic precipitation.
